Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
drafting the k-nearest neighbor averaging

\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
\begin{itemize}
\item Provide a detailed explanation of each neighbor-based method:
\begin{itemize}
\item \textbf{Independent learners:} Emphasize its simplicity and scalability but highlight its vulnerability to non-stationarity.
\item \textbf{NN-Averaging:} Describe the weight averaging process, discuss the influence of neighbor selection and averaging frequency, and mention weighted averaging variations.
\item \textbf{NN-Consensus:} Explain the selection of the best-performing network and its potential benefits and drawbacks.
\item \textbf{Experience Sharing:} Detail how agents share experiences, discuss asynchronous vs. synchronous sharing, and highlight its advantages in sparse interaction environments.
\end{itemize}
\end{itemize}
\section{Experimental Setup}\label{sec:experiments}
\begin{itemize}
\item Introduce the custom "collect the items" multi-agent environment:
\begin{itemize}
\item Describe the environment's dynamics, agent goals, and challenges.
\item Specify the action space (continuous and discrete) and observation space design.
\item Explain the reward structure, emphasizing the components that encourage cooperation.
\end{itemize}
\item Detail the implementation framework (Gymnasium, RLlib, PyTorch).
\item Specify the chosen RL algorithm (DQN) and its hyperparameter configuration.
\item Define the evaluation metrics: average episode length and average reward.
\end{itemize}

>>>>

\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning. 
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on deep Q-learning as the underlying learning algorithm,
and we consider the following strategies:
 k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.

Since these strategies are in general based on the notion of k-nearest neighbors,
we first introduce this concept and then describe each strategy in detail.
Formally, consider a \emph{weighted} communication graph $\mathcal{G} = (N, E, W)$, where $N$ is the set of agents, where $W: E \rightarrow \mathbb{R}^+$ is a weight function assigning a positive weight to each edge.  
Let $d(i,j)$ denote the shortest path length between agents $i$ and $j$ in the weighted graph $\mathcal{G}$. 
This can be calculated, for instance, using Dijkstra's algorithm. 
The length of a path is the sum of the weights of the edges along the path. If no path exists between $i$ and $j$, we set $d(i,j) = \infty$.
%
Then, the set of $k$-nearest neighbors of agent $i$ is defined as:
$$ \mathcal{N}_k(i) = { j \in N \setminus {i} \mid \text{rank}(d(i,j)) \leq k } $$
where $\text{rank}(d(i,j))$ denotes the rank of the distance $d(i,j)$ among all distances from agent $i$ to other agents in $N \setminus {i}$, sorted in ascending order (e..g, $\text{rank}(d(i,j)) = 1$ for the smallest distance).
Ties can be broken arbitrarily.

\subsection{K-Nearest Neighbor Averaging}
This approach averages an agent's Q-values with those of its nearest neighbors. 
Each agent maintains a local Q-network and shares its Q-values with its neighbors. 
The neighbors average the received Q-values and use the result to update their own Q-networks. 
This leverages the collective knowledge of neighboring agents to improve learning.
%
Formally, giving $\theta^t_i, i \in N$ the parameters of the Q-networks of the agent $i$ at the time $t$,
the update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \theta^{t+1}_i = \theta^t_i + \alpha \sum_{j \in \mathcal{N}_k(i)} \frac{1}{k} \theta^t_j
\end{equation}

\subsection{Nearest Neighboring Consensus}

\subsection{Experience Sharing}

\begin{itemize}
\item Provide a detailed explanation of each neighbor-based method:
\begin{itemize}
\item \textbf{Independent learners:} Emphasize its simplicity and scalability but highlight its vulnerability to non-stationarity.
\item \textbf{NN-Averaging:} Describe the weight averaging process, discuss the influence of neighbor selection and averaging frequency, and mention weighted averaging variations.
\item \textbf{NN-Consensus:} Explain the selection of the best-performing network and its potential benefits and drawbacks.
\item \textbf{Experience Sharing:} Detail how agents share experiences, discuss asynchronous vs. synchronous sharing, and highlight its advantages in sparse interaction environments.
\end{itemize}
\end{itemize}
\section{Experimental Setup}\label{sec:experiments}
\begin{itemize}
\item Introduce the custom "collect the items" multi-agent environment:
\begin{itemize}
\item Describe the environment's dynamics, agent goals, and challenges.
\item Specify the action space (continuous and discrete) and observation space design.
\item Explain the reward structure, emphasizing the components that encourage cooperation.
\end{itemize}
\item Detail the implementation framework (Gymnasium, RLlib, PyTorch).
\item Specify the chosen RL algorithm (DQN) and its hyperparameter configuration.
\item Define the evaluation metrics: average episode length and average reward.
\end{itemize}
